---
title: "map-statistics"
format: html
---

# Remove environment

```{r}
rm(list = ls())
gc()  
```

# Read in libraries

```{r}
library(conflicted)
library(terra)
library(tidyverse)
library(ggplot2)
library(viridis)

conflict_prefer("extract", "terra")
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
conflict_prefer("select", "dplyr")
```

# Read in coverage map, slope, mask, and digital elevation model

```{r}
dem <- terra::rast("raw_data/ortho/dem.tif")

CRS <- crs(dem)

coverage <- terra::rast('clean_data/ortho/orthomosaic_composition.tif') |>
  subset("sand") |>
  terra::project(CRS)


if(!file.exists('outputs/ortho/slope.tif')){
  slope <- terra::terrain(dem, v="slope", unit="degrees") |>
    terra::project(CRS)
  
  terra::writeRaster(slope, 'outputs/ortho/slope.tif')
} else {
  slope <- terra::rast('outputs/ortho/slope.tif') |>
    terra::project(CRS)
}


mask <- terra::vect('clean_data/ortho/grass-mask/mask_grass_paper.shp') |>
  terra::project(CRS)
```

# Plot mask over orthos

```{r}
create_plot <- function(raster, title, mask = NULL, color_scale = viridis::viridis) {
  rast <- raster
  plot(rast, main = title, col = color_scale(100))
  
  if (!is.null(mask)) {
    lines(mask, col = "red", lwd = 2)
  }
}

create_plot(dem, "Digital Elevation Model with Mask", mask = mask)
create_plot(coverage, "Vegetation Coverage with Mask", mask = mask)
create_plot(slope, "Slope with Mask", mask = mask, color_scale = terrain.colors)
```

## Create data

Resample the coverage map onto the slope grid so the two rasters share an identical
extent/resolution/CRS.

```{r}
coverage_resampled <- terra::resample(coverage, slope, method="bilinear")
names(coverage_resampled) <- "sand"
names(slope) <- "slope"
```

# Visual check

Mask the aligned rasters directly. (Previously this rebuilt the rasters from a data-frame of
xyz cell centres via `terra::rast(data, crs = CRS)`; inferring the grid from floating-point
coordinates produced a slightly irregular resolution — x and y differed in the ~10th decimal —
which made rasterising the mask polygon behave inconsistently, so the mask "worked sporadically".
Masking the already-registered rasters removes the round-trip and is stable.)

```{r}
sand_rast  <- terra::mask(coverage_resampled, mask, inverse = TRUE)
slope_rast <- terra::mask(slope,              mask, inverse = TRUE)

create_plot(sand_rast, "Sand", mask=mask)
create_plot(slope_rast, "Slope", mask=mask)
```


## Spearmans Rank Correlation 

```{r}
df <- c(sand_rast, slope_rast) |>
  terra::as.data.frame(xy=TRUE) |>
  as_tibble()

spearman_cor <- df |>
  select(-c(x,y)) |> 
  cor( method = "spearman")
```


